Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2020-12-15 jss/EpiCompare-paper/EpiCompare-paper.Rmd
tidy_dist_mat and calculation of new observation's depth / psuedo-densities relative to another set of observations (part 4): pkgdown creation

>>>>

Although epidemics are by definition a process that evolves over time, epidemics, including the recent COVID-19, often need to be compared in a time-invariant way to understand the processes at play. Additionally, many tools to examine the quantity of the population in each state along the infection process (for example: quantity of suspectible vs infected vs recovered individuals) don't always as intelligently capture the natural connections between the proportion of individuals in these states. Tools in \pkg{EpiCompare} attempt to give the user the ability to extend their toolkit to evaluate epidemics to also include time-invariant approaches and in this section we present benefits of the time-invariant analysis, with 1) motivation for time-invariant analysis through $R_0$ (the initial reproduction number), 2) time-invariant visualization tools for 3-state models (e.g. SIR models), and 3) potential for similar analysis for models with more states (e.g. SEIR models). \ben{Will need to rewrite this section "introduction".}

## Motivation through $R_0$ {short-title="r0" #r0}

$R_0$, the initial reproduction number, has been called the ``most important quantity in epdiemiology'' \citep[][]{Gallagher2020}. To see the importance of $R_0$, one need only read the newspaper \citep{Fisher2020} or look at the length of the table of estimated $R_0$ quantities for COVID-19 \citep{Aronson2020}. $R_0$ is also, maybe, the most famous \textit{time-invariant} numerical summary of an epidemic, and is commonly associated with the Susceptible-Infectious-Recovered (SIR) data/models. The SIR framework, composed of data and statistical models where individuals pass from Susceptible to Infectious to Recovered states is a common, if not the most common, modeling framework in infectious disease epidemiology (seen in examples of recent published works in \cite{FILLIN}).

In regards to traditional $X$ vs. time plots, $R_0$ is difficult to visualize. For example, in a SIR simulation with two models have the same value of $R_0$, one can see in Fig. \ref{fig:different-scales-standard} that there is no way of knowing that from looking at the graph (this figure will be commented on more below).

\begin{figure}[!ht]
%% Code to reproduce is in inst/paper_figs/different-time-scales.R
    \centering
    \includegraphics[width = .5\textwidth]{images/diff-time-standard.pdf}%\includegraphics[width = .5\textwidth]{images/diff-time-ternary.pdf}
    \caption{Bivariate view of \# in each state vs. time. Both data sets are generated using the same value of $R_0 = 2.8$ but have different values of $\beta$ and $\gamma$.\ben{We probably need to define $\beta/\gamma$.}
\ben{NEW: recommendation - make this plot actually have 3 examples (the current one), one where we have an affine transformation of the time scale, and one with the halleolof data.} \ben{Also these figures should be reproducible.}}
    \label{fig:different-scales-standard}
\end{figure}

## Ternary and time-invariant visualizations for $R_0$ {short-title="ternary visualizations" #sec:ternary}

It is possible to identify, at a glance, which SIR epidemic paths, like those in Fig. \ref{fig:different-scales-tern}  have the same value of $R_0$ via ``time-invariant'' visualizations, under some circumstances. Specifically, for SIR models we propose using ternary to examine an epidemic's trajectory in a more time-invariant manner.

Ternary plots are sometimes used in chemistry \citep[][]{Gillespie1976} but only rarely seen in the field of infectious disease epidemiology (to our knowledge, only seen in \cite{}).  Ternary plots illustrate the relationship of constrained 3D data, namely of points $(a_i, b_i, c_i) \in [0,1] \times [0,1] \times [0,1]$ with $a_i+ b_i + c_i =1$ and as such, are situational.  The SIR framework happens to be of the situational format required for ternary plots where $(S(t), I(t), R(t))$ are the number of susceptible, infectious, and recovered individuals in each state at time $t$ and $S(t) + I(t) + R(t) = N(t)$ where $N(t)$ is the population size at time $t$.

Ternary plots are time-invariant in the sense that the temporal scale does not explicitly apear in the visualization and as a consequence, allow for direct comparisons of outbreaks on different time scales (e.g. days vs. years) or of SIR data of the same disease in different areas (our first case study in Sec. \ref{sec:ex1} focuses on this).

To show that ternary plots have to potential to aid in the comparison of different epidemic's $R_0$s -in certain circumstances- we first need to introduce the classic, deterministic Kermack and McKendrick \cite{CITETHIS?} SIR model where the initial number in each state $(S(0), I(0), R(0))$ are known and $N$ the population size is constant.  The movement of individuals among the states is given by the following differential equations (Eq. \eqref{eq:sir-ode}) where $\beta$ is the average infection rate and $\gamma$ is the average recovery rate,
\begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align}

\ben{SUGGESTION: define $R_0$ mathematically here (though referred to in figure 2 - which still needs to be defined there).}

If we have two SIR models that follow the Kermack and McKendrick SIR model, have the same percentage of individuals in the initial states, and the same value of $R_0$ then the two models will create exactly overlapping paths in a ternary plot.

More formally, let two Kermack and McKendrick (see Eq. \eqref{eq:sir-ode}) SIR models be denoted $(S_1(t), I_1(t), R_1(t))$ and $(S_2(t), I_2(t), R_2(t))$, respectively, for $t > 0$. Assume both models have initial values $(S(0), I(0), R(0))$.  Let $R_0 = \frac{\beta_1}{\gamma_1} = \frac{\beta_2}{\gamma_2}$ where $\beta_i$ and $\gamma_i$ are the average infection rate and recovery rate, respectively, for SIR model $i=1, 2$.  Equivalently, $\beta_2 = a \beta_1$ if and only if $\gamma_2 = a \gamma_1$ for some $a > 0$.

\begin{theorem}\label{thm:sir-scale}
Let there be two SIR models as described above.  Then for all $t > 0$ there exists an $s$ such that $(S_1(t), I_1(t), R_1(t)) = (S_2(s), I_2(s), R_2(s))$.  Moreover, $s = \frac{1}{a}t$.
\end{theorem}

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent result from \cite{Harko2014} and is shown in detail in Appendix \ref{app:proof}.  The consequence of Theorem \ref{thm:sir-scale} is that for two SIR models that have the same initial percent of individuals in each state and $R_0$ then for every point on the epidemic path of the first SIR model is also a point on the epidemic path of the second SIR model. Taking the sample simulations from Fig. \ref{fig:different-scales-standard}, Fig. \ref{fig:different-scales-tern} presents these models in a ternary plot.

\begin{figure}[!ht]
%% Code to reproduce is in inst/paper_figs/different-time-scales.R
    \centering
    \includegraphics[width = .5\textwidth]{images/diff-time-ternary.pdf}
    \caption{Ternary view of \# in each state.  Both this Fig. and Fig. \ref{fig:different-scales-standard} display the same two sets of data.  Both data sets are generated using the same value of $R_0 = 2.8$ but have different values of $\beta$ and $\gamma$. \ben{We probably need to define $\beta/\gamma$.}  While there are obvious are differences in Fig.\ref{fig:different-scale-standard}, the data sets look quite similar in the ternary view. \ben{I might suggest using opacity to show time scale here?} \ben{Make this reproducible too?}}
    \label{fig:different-scales-tern}
\end{figure}

In \pkg{EpiCompare}, we also provide visualization tools to aid in comparing models in the ternary / time-invariant space, which is presented in more details in Section \ref{sec:time-invariant-in-simplex}.

## Simplexes beyond ternary plots and time-invariant comparison tools {short-title="Simplexes beyond ternary plots" #sec:time-invariant-in-simplex}

Beyond ternary plots, and the associated SIR models, we can extend the time-invariant approach described in Section \ref{sec:ternary} to higher dimensions (i.e. models with more states). The constraints in 3d that are met with the SIR model (that is $\sum_{i=1}^3 (\text{number in state}(i)) = N(t)$) actually represents a spaces of 3d simplexes, and the ternary plot specifically represents these after scaling to examine such values as proportions (ternary plots are known as a 3d unit simplex due to it's scaling). This same scaling for larger models (i.e. with more states) can be done onto different simplexes. In this package we present tools to help compare models (mostly through similations), and this tool compares these objects after projecting them into a one-dimension less space through the simplexical structure of the data.

In \pkg{EpiCompare} we allow for the compare epidemics in these higher dimensional spaces by first projecting onto these simplexes. Even though higher dimensional models may not be able to visualized, we provide multiple tools to aid in the comparison of models and epidemics. The first of which uses multiple simulations under specific model parameters to assess the bairabilty of the model fit. In \pkg{EpiCompare} we provide ways to create prediction regions for a true epidemic under the model assumptions using these simulations. These regions require representing multi-dimensional structures for functions to completely contain epidemics, and treat these simulations and epidmeics as filamental objects. We extend off of papers like \citet{Dalmasso2019a} to create these bands. These high dimensional bands allow the user to assess if the true epidemic is within the band (there-by assessing the model's representation of the epidemic), and also compare who different models are from each other through distances that compare sets. We recommend using the Hausdorff distance to compare such sets as it captures how much bigger the sets would have to expand to cover each other, and is defined mathematically as 
\[
d_\text{Hausdorff}(S_1, S_2) = \max \left\{ \sup_{x \in S_1} \inf_{y \in S_2} d(x,y) \;,\; \sup_{y \in S_2} \inf_{x \in S_1} d(x,y)\right\}\;.
\]

### 2020-12-15 jss/EpiCompare-paper/EpiCompare-paper.tex
tidy_dist_mat and calculation of new observation's depth / psuedo-densities relative to another set of observations (part 4): pkgdown creation

% Pandoc citation processing

@@

The current epidemic of COVID-19 also highlights that infectious disease
models are only one piece of the overall analysis pipeline. University
based resources like John Hopkin's and government numerical dashboards
(across all levels of government) during the COVID-19 epidemic remind us
that descriptive statistics and visualization can be a important first
step in the process (multiple \cite{}?).\} Still, rightly so, a large
amount of theoretical work goes into modeling epidemics, with different
models focusing at the individual / agent level, network structure or
just aggregate flows (review paper \cite{}?). All placing individuals /
proportions of the populations into different states (e.g.~suspectible,
exposed, infected, recovered, etc.). With all these models, review and
comparison papers in the literature and through MIDAS (Models of
Infectious Disease Agent Study) Control Center helps the individual
practitioner decide the correct approach.
\textcolor{orange}{along with expertise from healthcare professionals...}

@@

This package also emphasizes the value of approach epidemics in a
\textit{time-invariant} way, not constrained to things like a specific
time scale, or initial date of the first infection. Although epidemics
are by definition a process that evolves over time, epidemics, including
the recent COVID-19, often need to be compared in a time-invariant way
to understand the processes at play. Additionally, many tools to examine
the quantity of the population in each state along the infection process
(for example: quantity of suspectible vs infected vs recovered
individuals) don't always as intelligently capture the natural
connections between the proportion of individuals in these states. Tools
in \pkg{EpiCompare} attempt to give the user the ability to extend their
toolkit to evaluate epidemics to also include time-invariant approaches.
The goal of \pkg{EpiCompare} is not to supplant existing infectious
disease modeling tools and software but, rather, is a concerted effort
to create standard and fair comparisons among models developed for
disease outbreaks and outbreak data.

@@

\hypertarget{visualize}{%
\subsection{Visualize}\label{visualize}}

One issue that the COVID-19 outbreak has brought to the forefront is the
importance of figures and graphs to inform, support, and even persuade
lay people of technical and scientific results \cite{}. Quantities and
trends shown in these visualizations include exponential growth curves,
S-curves, peak day and percent of infectious indviduals, ICU capacity,
and deaths over time and uncertainty of such estimates.

\pkg{EpiCompare} supports traditional visualizations (e.g.~\# in
condition vs.~time) and also introduces a novel visualization of
epidemics via time free analysis. In particular, we use ternary plots
that show Susceptible-Infectious-Recovered data/models (SIR). The
advantages of these ternary plots are 1) a standard way to compare
epidemics with different outbreak times and time scales, 2) a novel
interpretation of \(R_0\), the initial reproduction number under some
circumstances, and 3) clear visualization of joint SIR uncertainty via
conformal bands.

Ternary plots are sometimes used in chemistry \citep{} but only rarely
seen in the field of infectious disease epidemiology (to our knowledge,
only seen in \cite{}). Ternary plots illustrate the relationship of
constrained 3D data, namely of points
\((a_i, b_i, c_i) \in [0,1] \times [0,1] \times [0,1]\) with
\(a_i+ b_i + c_i =1\) and as such, are situational. The SIR framework,
composed of data and statistical models where individuals pass from
Susceptible to Infectious to Recovered states is a common, if not the
most common, modeling framework in infectious disease epidemiology (seen
in examples of recent published works in \cite{}), and also happens to
be of the situational format required for ternary plots where
\((S(t), I(t), R(t))\) are the number of susceptible, infectious, and
recovered individuals in each state at time \(t\) and
\(S(t) + I(t) + R(t) = N(t)\) where \(N(t)\) is the population size at
time \(t\).

Ternary plots are time-free in the sense that the temporal scale does
not explicitly apear in the visualization and as a consequence, allow
for direct comparisons of outbreaks on different time scales (e.g.~days
vs.~years) or of SIR data of the same disease in different areas (our
first case study in Sec. \ref{sec:ex1} focuses on this).

Ternary plots also, under certain conditions, have a special
relationship with \(R_0\), the initial reproduction number, which has
been called the ``most important quantity in epidemiology'\,' \citep{}.
The reproduction is itself a time-free concept and is the average number
of secondary infections primary infection will generate when the primary
infection is added to a completely susceptible population. To see the
importance of \(R_0\), one need only read the newspaper \citep{} or look
at the length of the table of estimated \(R_0\) quantities for COVID-19
\citep{}.

In regards to traditional \(X\) vs.~time plots, \(R_0\) is difficult to
visualize. For example, as seen in Fig. \ref{fig:r0-trad}, it is
possible that two models have the same value of \(R_0\) but there is no
way of knowing that from looking at the graph.

It is possible to identify, at a glance, which epidemic paths have the
same value of \(R_0\) via ternary plot visualizations, under some
circumstances. To show, this we first need to introduce the classic,
deterministic Kermack and McKendrick SIR model where the initial number
in each state \((S(0), I(0), R(0))\) are known and \(N\) the population
size is constant. The movement of individuals among the states is given
by the following differential equations (Eq. \eqref{eq:sir-ode}) where
\(\beta\) is the average infection rate and \(\gamma\) is the average
recovery rate, \begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align}

@@

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent
result from \cite{} and is shown in detail in Appendix \ref{app:proof}.
The consequence of Theorem \ref{thm:sir-scale} is that for two SIR
models that have the same initial percent of individuals in each state
and \(R_0\) then for every point on the epidemic path of the first SIR
model is also a point on the epidemic path of the second SIR model.

>>>>

@@

The current epidemic of COVID-19 also highlights that infectious disease
models are only one piece of the overall analysis pipeline. University
based resources like John Hopkin's and government numerical dashboards
(across all levels of government) during the COVID-19 epidemic remind us
that descriptive statistics and visualization can be a important first
step in the process (multiple \cite{}?).\}
\textcolor{orange}{add comment about NYT or something} Still, rightly
so, a large amount of theoretical work goes into modeling epidemics,
with different models focusing at the individual / agent level, network
structure or just aggregate flows (review paper \cite{}?). All placing
individuals / proportions of the populations into different states
(e.g.~suspectible, exposed, infected, recovered, etc.). With all these
models, review and comparison papers in the literature and through MIDAS
(Models of Infectious Disease Agent Study) Control Center helps the
individual practitioner decide the correct approach.
\textcolor{orange}{along with expertise from healthcare professionals...}

@@

This package also emphasizes the value of approach epidemics in a
\textit{time-invariant} way, not constrained to things like a specific
time scale, or initial date of the first infection. Although epidemics
are by definition a process that evolves over time, epidemics, including
the recent COVID-19, often need to be compared in a time-invariant way
to understand the processes at play. Additionally, many tools to examine
the quantity of the population in each state along the infection process
(for example: quantity of suspectible vs infected vs recovered
individuals) don't always as intelligently capture the natural
connections between the proportion of individuals in these states. Tools
in \pkg{EpiCompare} attempt to give the user the ability to extend their
toolkit to evaluate epidemics to also include time-invariant approaches.
The goal of \pkg{EpiCompare} is not to supplant existing infectious
disease modeling tools and software but, rather, is a concerted effort
to create standard and fair comparisons among models developed for
disease outbreaks and outbreak data.
\textcolor{orange}{flesh out time invariance introduction to better explain to non-experts what we mean.  perhaps mention R0}

@@

Although epidemics are by definition a process that evolves over time,
epidemics, including the recent COVID-19, often need to be compared in a
time-invariant way to understand the processes at play. Additionally,
many tools to examine the quantity of the population in each state along
the infection process (for example: quantity of suspectible vs infected
vs recovered individuals) don't always as intelligently capture the
natural connections between the proportion of individuals in these
states. Tools in \pkg{EpiCompare} attempt to give the user the ability
to extend their toolkit to evaluate epidemics to also include
time-invariant approaches and in this section we present benefits of the
time-invariant analysis, with 1) motivation for time-invariant analysis
through \(R_0\) (the initial reproduction number), 2) time-invariant
visualization tools for 3-state models (e.g.~SIR models), and 3)
potential for similar analysis for models with more states (e.g.~SEIR
models).
\textcolor{violet}{Will need to rewrite this section "introduction".}

\subsection[r0]{Motivation through \(R_0\)}\label{r0}

\(R_0\), the initial reproduction number, has been called the ``most
important quantity in epdiemiology'' \citep[][]{Gallagher2020}. To see
the importance of \(R_0\), one need only read the newspaper
\citep{Fisher2020} or look at the length of the table of estimated
\(R_0\) quantities for COVID-19 \citep{Aronson2020}. \(R_0\) is also,
maybe, the most famous \textit{time-invariant} numerical summary of an
epidemic, and is commonly associated with the
Susceptible-Infectious-Recovered (SIR) data/models. The SIR framework,
composed of data and statistical models where individuals pass from
Susceptible to Infectious to Recovered states is a common, if not the
most common, modeling framework in infectious disease epidemiology (seen
in examples of recent published works in \cite{FILLIN}).

In regards to traditional \(X\) vs.~time plots, \(R_0\) is difficult to
visualize. For example, in a SIR simulation with two models have the
same value of \(R_0\), one can see in Fig.
\ref{fig:different-scales-standard} that there is no way of knowing that
from looking at the graph (this figure will be commented on more below).

\begin{figure}[!ht]
%% Code to reproduce is in inst/paper_figs/different-time-scales.R
    \centering
    \includegraphics[width = .5\textwidth]{images/diff-time-standard.pdf}%\includegraphics[width = .5\textwidth]{images/diff-time-ternary.pdf}
    \caption{Bivariate view of \# in each state vs. time. Both data sets are generated using the same value of $R_0 = 2.8$ but have different values of $\beta$ and $\gamma$.\textcolor{violet}{We probably need to define $\beta/\gamma$.}
\textcolor{violet}{NEW: recommendation - make this plot actually have 3 examples (the current one), one where we have an affine transformation of the time scale, and one with the halleolof data.} \textcolor{violet}{Also these figures should be reproducible.}}
    \label{fig:different-scales-standard}
\end{figure}

\subsection[ternary visualizations]{Ternary and time-invariant
visualizations for \(R_0\)}\label{sec:ternary}

It is possible to identify, at a glance, which SIR epidemic paths, like
those in Fig. \ref{fig:different-scales-tern} have the same value of
\(R_0\) via ``time-invariant'' visualizations, under some circumstances.
Specifically, for SIR models we propose using ternary to examine an
epidemic's trajectory in a more time-invariant manner.

Ternary plots are sometimes used in chemistry \citep[][]{Gillespie1976}
but only rarely seen in the field of infectious disease epidemiology (to
our knowledge, only seen in \cite{}). Ternary plots illustrate the
relationship of constrained 3D data, namely of points
\((a_i, b_i, c_i) \in [0,1] \times [0,1] \times [0,1]\) with
\(a_i+ b_i + c_i =1\) and as such, are situational. The SIR framework
happens to be of the situational format required for ternary plots where
\((S(t), I(t), R(t))\) are the number of susceptible, infectious, and
recovered individuals in each state at time \(t\) and
\(S(t) + I(t) + R(t) = N(t)\) where \(N(t)\) is the population size at
time \(t\).

Ternary plots are time-invariant in the sense that the temporal scale
does not explicitly apear in the visualization and as a consequence,
allow for direct comparisons of outbreaks on different time scales
(e.g.~days vs.~years) or of SIR data of the same disease in different
areas (our first case study in Sec. \ref{sec:ex1} focuses on this).

To show that ternary plots have to potential to aid in the comparison of
different epidemic's \(R_0\)s -in certain circumstances- we first need
to introduce the classic, deterministic Kermack and McKendrick
\cite{CITETHIS?} SIR model where the initial number in each state
\((S(0), I(0), R(0))\) are known and \(N\) the population size is
constant. The movement of individuals among the states is given by the
following differential equations (Eq. \eqref{eq:sir-ode}) where
\(\beta\) is the average infection rate and \(\gamma\) is the average
recovery rate, \begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align}

\textcolor{violet}{SUGGESTION: define $R_0$ mathematically here (though referred to in figure 2 - which still needs to be defined there).}

@@

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent
result from \cite{Harko2014} and is shown in detail in Appendix
\ref{app:proof}. The consequence of Theorem \ref{thm:sir-scale} is that
for two SIR models that have the same initial percent of individuals in
each state and \(R_0\) then for every point on the epidemic path of the
first SIR model is also a point on the epidemic path of the second SIR
model. Taking the sample simulations from Fig.
\ref{fig:different-scales-standard}, Fig.
\ref{fig:different-scales-tern} presents these models in a ternary plot.

\begin{figure}[!ht]
%% Code to reproduce is in inst/paper_figs/different-time-scales.R
    \centering
    \includegraphics[width = .5\textwidth]{images/diff-time-ternary.pdf}
    \caption{Ternary view of \# in each state.  Both this Fig. and Fig. \ref{fig:different-scales-standard} display the same two sets of data.  Both data sets are generated using the same value of $R_0 = 2.8$ but have different values of $\beta$ and $\gamma$. \textcolor{violet}{We probably need to define $\beta/\gamma$.}  While there are obvious are differences in Fig.\ref{fig:different-scale-standard}, the data sets look quite similar in the ternary view. \textcolor{violet}{I might suggest using opacity to show time scale here?} \textcolor{violet}{Make this reproducible too?}}
    \label{fig:different-scales-tern}
\end{figure}

In \pkg{EpiCompare}, we also provide visualization tools to aid in
comparing models in the ternary / time-invariant space, which is
presented in more details in Section
\ref{sec:time-invariant-in-simplex}.

\subsection[Simplexes beyond ternary plots]{Simplexes beyond ternary
plots and time-invariant comparison
tools}\label{sec:time-invariant-in-simplex}

Beyond ternary plots, and the associated SIR models, we can extend the
time-invariant approach described in Section \ref{sec:ternary} to higher
dimensions (i.e.~models with more states). The constraints in 3d that
are met with the SIR model (that is
\(\sum_{i=1}^3 (\text{number in state}(i)) = N(t)\)) actually represents
a spaces of 3d simplexes, and the ternary plot specifically represents
these after scaling to examine such values as proportions (ternary plots
are known as a 3d unit simplex due to it's scaling). This same scaling
for larger models (i.e.~with more states) can be done onto different
simplexes. In this package we present tools to help compare models
(mostly through similations), and this tool compares these objects after
projecting them into a one-dimension less space through the simplexical
structure of the data.

In \pkg{EpiCompare} we allow for the compare epidemics in these higher
dimensional spaces by first projecting onto these simplexes. Even though
higher dimensional models may not be able to visualized, we provide
multiple tools to aid in the comparison of models and epidemics. The
first of which uses multiple simulations under specific model parameters
to assess the bairabilty of the model fit. In \pkg{EpiCompare} we
provide ways to create prediction regions for a true epidemic under the
model assumptions using these simulations. These regions require
representing multi-dimensional structures for functions to completely
contain epidemics, and treat these simulations and epidmeics as
filamental objects. We extend off of papers like \citet{Dalmasso2019a}
to create these bands. These high dimensional bands allow the user to
assess if the true epidemic is within the band (there-by assessing the
model's representation of the epidemic), and also compare who different
models are from each other through distances that compare sets. We
recommend using the Hausdorff distance to compare such sets as it
captures how much bigger the sets would have to expand to cover each
other, and is defined mathematically as \[
d_\text{Hausdorff}(S_1, S_2) = \max \left\{ \sup_{x \in S_1} \inf_{y \in S_2} d(x,y) \;,\; \sup_{y \in S_2} \inf_{x \in S_1} d(x,y)\right\}\;.
\]
